---
title: "R Notebook"
output: html_notebook
---

Plots for host and env alpha diversity models

```{r}
rm(list=ls())
```

```{r}
#install.packages("interactions")
library(jtools)
library(interactions)
library(ggplot2)
library(MuMIn)
library(lme4)
library(microbiome)
library(cowplot)
library(ggpubr)

```

https://stackoverflow.com/questions/28633537/partial-residual-plot-based-on-model-average-coefficients-in-r
https://stackoverflow.com/questions/25695565/predict-with-arbitrary-coefficients-in-r/25697144#25697144
https://stackoverflow.com/questions/56924967/convert-model-averaging-results-from-mumin-to-plot-in-visreg-or-effects-packages
https://www.researchgate.net/post/How_can_i_plot_Estimates_of_Average_Models_using_Full_Average_values


Difficult to get plotting packages working with model averaged results. Is a hack to take the full model or one of the most complex 'top.model' and to replace the coefficients with those of the model.avg object? 

The habitat variable is centred by the arm::standardize command and so is represented by numeric values rather than categorical factor designations. 
* where -0.79 is conifer and 0.21 is deciduous
* maybe i could replace those values with strings before plotting?
* Or just reference those values and overwrite the labels in the plot
* really good blog post which deals with plotting centred predictors with interactions: https://seananderson.ca/2014/07/30/centering-interactions/

###########################################################################################################################################
###########################################################################################################################################
###########################################################################################################################################
# read in data
```{r}
model.set.all.shannon <- readRDS("../Data/dredgeSaves/model.set.all.shannon")
model.set.all.chao <- readRDS("../Data/dredgeSaves/model.set.all.chao")

top.models <- get.models(model.set.all.shannon, subset = delta<2, REML = TRUE)
top.models2 <- get.models(model.set.all.chao, subset = delta<2, REML = TRUE)

# model.set.adult.shannon <- readRDS("../Data/dredgeSaves/model.set.adult.shannon")
# model.set.adult.chao <- readRDS("../Data/dredgeSaves/model.set.adult.chao")
# 
# top.models.3 <- get.models(model.set.adult.shannon, subset = delta<2, REML = TRUE)
# top.models.4<- get.models(model.set.adult.chao, subset = delta<2, REML = TRUE)
```

```{r}
phylo.spring <- readRDS(file = "../Data/phylo-spring.rds")
metadata <- meta(phylo.spring) # doesnt maintain Date data-type
```

# All birds
## Distance to edge, shannon

"CUD Bright" from jtools and ultimately Paul Tol is color scheme.
broodSizeXage.proteo.all[["plot_env"]][["colors"]]
   Day 8    Day 15     Adult 
"#49b7fc" "#ff7b00" "#17d898" 
```{r}
#define colors for each point
#pnt.color <- ifelse(metadata$ageBinned=="adult","#17d898" , ifelse(metadata$ageBinned=="2week","#ff7b00" ,"#49b7fc"))

#scales::hue_pal(broodSizeXage.proteo.all)
```

```{r}
all.sh.lmm.top <- top.models[[1]]

distance.shannon.all<-effect_plot(all.sh.lmm.top,pred=z.DistanceToEdge,x.label="Distance to edge", y.label="log(Shannon)", interval =TRUE, int.type = "confidence", partial.residuals = TRUE, data = all.sh.lmm.top@frame) #point.color = pnt.color,

distance.shannon.all
```

## Distance to edge, chao
```{r}
all.chao.lmm.top <- top.models2[[1]]
summary(all.chao.lmm.top)

#all.chao.lmm.top@frame$c.habitat <- ifelse(all.chao.lmm.top@frame$c.habitat<0, "conifer", "deciduous")

distance.chao.all<-effect_plot(all.chao.lmm.top, pred=z.DistanceToEdge, x.label="Distance to edge", y.label="log(Chao)", interval =TRUE, int.type = "confidence", partial.residuals = TRUE,  data = all.chao.lmm.top@frame) #point.color = pnt.color,

distance.chao.all
```
## Distance x age, chao
NOTE: using second top ranked model for plot because this contains the age x distance interaction

Slopes for the age x distance interaction should be negative according to model estimates but slopes in plot are positive
* is the slope calculated as the slope for distance (0.47) plus the interaction coefficient (-0.46)? Instead of what im expecting distance plus ageBinned1 plus interaction?
*note: intercept refers to grand mean not reference level mean and the coefficients are main effects rather than marginal effects, so that distance coefficient refers the main effect of distance across age groups. The interaction slopes are therefore relative to an unseen D8 slope and dont disagree with the model outputs.
```{r}
all.chao.lmm.top <- top.models2[[2]]
summary(all.chao.lmm.top)

#all.chao.lmm.top@frame$c.habitat <- ifelse(all.chao.lmm.top@frame$c.habitat<0, "conifer", "deciduous")

distance.chao.age <- interact_plot(all.chao.lmm.top, pred=z.DistanceToEdge, modx=ageBinned, x.label="Distance to edge", y.label="log(Chao)", interval =TRUE, int.type = "confidence", partial.residuals = TRUE, legend.main = "Age", modx.labels = c("Day 8", "Day 15", "Adult"), data = all.chao.lmm.top@frame) #point.color = pnt.color,

distance.chao.age
```

## Age, chao
Age is significant in full averaged model while age*habitat is only included 4th top model

```{r}
all.chao.lmm.top <- top.models2[[1]]
summary(all.chao.lmm.top)
#all.chao.lmm.top@frame$c.habitat <- ifelse(all.chao.lmm.top@frame$c.habitat<0, "conifer", "deciduous")

age.chao.all<-effect_plot(all.chao.lmm.top, pred=ageBinned, x.label="Age", y.label="log(Chao)", interval =TRUE, int.type = "confidence", partial.residuals = TRUE, data = all.chao.lmm.top@frame)

age.chao.all
```

## age*habitat, chao
Age is significant in full averaged model while age*habitat is only included 4th top model
```{r}
all.chao.lmm.top <- top.models2[[4]]
summary(all.chao.lmm.top)

all.chao.lmm.top@frame$c.habitat <- ifelse(all.chao.lmm.top@frame$c.habitat<0, "Conifer", "Deciduous")

ageXhabitat.chao.all<-cat_plot(all.chao.lmm.top, pred="ageBinned", modx="c.habitat", x.label="Age", y.label="log(Chao)", legend.main = "Habitat", pred.labels = c("Day 8", "Day 15", "Adult"), interval =TRUE, int.type = "confidence", partial.residuals = TRUE, jitter = 0.4, point.alpha = 0.3, data = all.chao.lmm.top@frame)

ageXhabitat.chao.all
```

Add significance indicators
* was getting error without specifying data
* add bracket to connect significantly different means and lines to indicate significantly different slopes/interactions
* note: can refer to factors (i.e. age levels) as numbers instead of specifying the string. So can use 1 instead of "1week"
```{r}
ageXhabitat.chao.all <- ageXhabitat.chao.all +
  geom_bracket(
    data = all.chao.lmm.top@frame,
    xmin = "1week", xmax = "2week", y.position = 8,
    label = "p = 0.02",
    colour = "red"
  ) + geom_segment(
    x = 1 + 0.25,
    y = 6.3,
    xend = 2 + 0.19,
    yend = 5.2,
    colour = "black"
  ) + geom_segment(
    x = 1 - 0.2,
    y = 5.85,
    xend = 2 - 0.25,
    yend = 5.5,
    colour = "black"
  ) + geom_bracket(
    data = all.chao.lmm.top@frame,
    xmin = 1 - 0.5, xmax = 2 + 0.5, y.position = 4,
    tip.length = c(-.02, -0.02),
    label = "p = 0.025",
    colour = "black"
  )

ageXhabitat.chao.all
```


# Combine plots

```{r}
theme_classic2 <- function(base_size = 12, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      legend.position = "none",
      panel.border     = element_blank(),
      axis.line        = element_line(colour = "black"),
      panel.grid.major=element_line(colour="grey", size=0.5, 3),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      legend.key       = element_blank()
    )
}
```

```{r}
#p4.legend <- get_legend(broodSizeXage.proteo.all) # from script 05c_

p1 <- cowplot::plot_grid(distance.shannon.all+theme_classic2(), distance.chao.all+theme_classic2(), nrow = 1,rel_widths = c(1,1,0.4)) #p4.legend
p1
```
```{r}
#p4.legend <- get_legend(broodSizeXage.proteo.all) # from script 05c_

p1.age <- cowplot::plot_grid(distance.shannon.all+theme_classic(), distance.chao.age+theme_classic(), nrow = 1,rel_widths = c(1,1.3,0.4), labels = c("A", "B")) #p4.legend

p1.age

ggsave("Fig2_distance_diversity.tiff", plot = p1.age, device = "tiff", path = "../R_figures", width = 17.5, height = 10, units = "cm", dpi = 300)
```


```{r}
ageXhabitat.chao.all.classic <- ageXhabitat.chao.all+theme_classic()

ageXhabitat.chao.all.classic
 
ggsave("Fig1_ageXhabitat.chao.all.tiff", plot = ageXhabitat.chao.all.classic, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 300)
```

```{r}
# # make grid of age plots
# p1 <- plot_grid(age.proteo.adult+theme_classic2()+ylab(NULL)+labs(title="Proteobacteria"), age.firmic.adult+theme_classic2()+ylab(NULL)+labs(title="Firmicutes")) 
# p1.legend <- get_legend(age.proteo.adult)
# 
# # fake legend to align properly with below plots
# p1.grid <- plot_grid(p1, p1.legend, rel_widths = c(3, .4), labels = c("A", "B", "") )
# 
# #make grid of all adult interaction plots
# p2 <- plot_grid(broodSizeXsex.proteo.adult+theme_classic2()+ylab(NULL), broodSizeXsex.firmic.adult+theme_classic2()+ylab(NULL), distanceXsex.proteo.adult+theme_classic2()+ylab(NULL), distanceXsex.firmic.adult+theme_classic2()+ylab(NULL), ncol = 2, labels = c("C", "D", "E", "F"), label_x = 0, label_y = 1.1)
# 
# # get common legend of all interaction plots
# p2.legend <- get_legend(broodSizeXsex.proteo.adult)
# 
# p2.grid <- plot_grid(p2, p2.legend, rel_widths = c(3, .4))
# 
# # combine
# adult.grid <- plot_grid(p1.grid, p2.grid, nrow = 2, rel_heights = c(1,2))
# 
# adult.grid
```
