---
title: "R Notebook"
output: html_notebook
---

Make repeatability plot

Plot D8 alpha diversity against D15 alpha diversity and colour by individual to determine whether low repeatability is driven by low between or high within variability.

Wilson 2018: Repeatability can be low because of (a) high within individual variation, or (b) low between individual variation.

```{r}
library(cowplot)
```

# get paired dataframe

```{r Extract paired sample data, echo=FALSE}
indiv.table <- table(nestling.metadata$bird.ID)
paired.table <- subset(indiv.table,indiv.table>1)
paired.IDs <- rownames(paired.table)

#paired.D8.shannon <- metadata[(metadata$bird.ID %in% paired.IDs) && (metadata$ageBinned=="1week"),"shannon"]

# subset df instead
paired.df <- subset(nestling.metadata, bird.ID %in% paired.IDs & (ageBinned=="1week"|ageBinned=="2week"))

# remove samples >2
#which(indiv.table>2) #CB60C=3, CB6NA, KB922B=3, KB924B=3, KB61D
remove.paired <- c("CB60-C-D9.S22", "CB6-16-VZ72614.S28", "KB61D-D12.S15", "KB916C-D9.S81","KB922B-D10.S84", "KB924-B-9-201.S89", "CB6-16-VZ72616.S29","CB6-D16-VZ72615.S57","CB6-D16-VZ72617.S58") 
#remove.paired <- c()
paired.df <- subset(paired.df, !BIOM.ID %in% remove.paired)

#indiv.table <- table(paired.df$bird.ID)
#which(indiv.table>2)

# subset paired samples and rename columns
paired.d8 <- subset(paired.df,ageBinned=="1week")
paired.d15 <- subset(paired.df,ageBinned=="2week")
```

## Df creation

```{r}
paired.matched <- merge(paired.d8[,c("BIOM.ID","ageBinned","nest","shannon","chao1","bird.ID","habitat", "ageDays", "Tarsus", "Weight", "wing", "broodSizeWhenSampled", "broodSizeMax", "Survive", "scaled.mass.chick", "site", "SequencePlate", "layDateFirst", "clutchSize")], paired.d15[,c("BIOM.ID","ageBinned","shannon","chao1","bird.ID","ageDays", "Tarsus", "Weight", "wing", "broodSizeWhenSampled", "scaled.mass.chick","clutchSize","SequencePlate", "DistanceToEdge")], by = "bird.ID")
# Tarsus and wing not measured for D8's?
```

Create variable: age difference in days between paired samples.
```{r}
paired.matched$ageDiff <- as.numeric(as.character(paired.matched$ageDays.y))-as.numeric(as.character(paired.matched$ageDays.x))

```

Create centred and scaled df. 
* Already done in 06_Repeatability script

```{r}
#dput(colnames(paired.matched))
# numeric.predictors <- c("shannon.x","shannon.y","chao1.x","chao1.y","Tarsus.x", "Weight.x", "wing.x", "broodSizeWhenSampled.x", "broodSizeMax", "scaled.mass.chick.x", "layDateFirst", "clutchSize.x", "Tarsus.y", "wing.y", "broodSizeWhenSampled.y", "scaled.mass.chick.y", "clutchSize.y", "DistanceToEdge","ageDiff")
# paired.matched.scaled <- paired.matched
# paired.matched.scaled[,numeric.predictors] <- scale(paired.matched[,numeric.predictors], scale = TRUE, center = TRUE)

```

```{r}
table(paired.df$ageBinned, paired.df$habitat)
```

# add RA info to paired.df
```{r}
paired.df <- merge(paired.df, proteobacteria.df[,c("BIOM.ID","Abundance")]) %>%
  rename("Proteobacteria.Abundance" = "Abundance")

paired.df <- merge(paired.df, firmicutes.df[,c("BIOM.ID","Abundance")]) %>%
  rename("Firmicutes.Abundance" = "Abundance")
```

# Add PC1 and PC2 to df
```{r}
paired.df <- merge(paired.df, df_out[,c("BIOM.ID", "PC1", "PC2")]) 
```

# Plotting

Make plot of D8 and D15 alpha diversity and connect individuals with lines
Maybe add plot of repeatability estimates from rtpR
Could colour lines according to site then nest then individual? or just have lines connecting individuals coloured by nest

Is this the best way of plotting repeatability? This is confusing me because how can individuals have 'consistent differences' when they are only measured twice and hence there is only a single slope for each individual?
  Is it that an individual at D8 retains a similar difference to another individual, even when they have both aged from D8 to D15?

Could try same plot as below but facet by positive or negative change

# Shannon 
## Reaction norm plot


```{r}
ggplot(paired.df, aes(ageBinned, shannon, fill=ageBinned)) +
  geom_boxplot() +
  geom_line(aes(group=bird.ID), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  theme(legend.position = "none")
```

```{r}
paired.boxplot.shannon <- ggplot(paired.df, aes(ageBinned, shannon, fill=ageBinned)) +
  geom_boxplot(aes(alpha = 0.1)) +
  geom_line(aes(group=bird.ID, colour = nest), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  xlab("Age") +
  ylab("Shannon diversity") +
  theme_classic() +
  theme(legend.position = "none")

paired.boxplot.shannon
```

## Plot w/ rptR estimate 

```{r}
df <- rpt.shannon.birdID.nest.site.fixed

df.rptr <- cbind(t(df[["R"]]),df[["CI_emp"]]) %>% 
  tibble::rownames_to_column(var = "variable") %>%
  rename("rpt" = "t(df[[\"R\"]])", "lower" = "2.5%", "upper" = "97.5%") # including escapes for \"
  
df.rptr$variable <- factor(df.rptr$variable, ordered = T, levels = c("site", "nest", "bird.ID", "Fixed"))

df.rptr

#str(df.rptr)
```
I should add something here to indicate significance. Though is this small LRT values or CI's that dont overlap zero? Sometimes these disagree e.g. for site
```{r}
rpt.confint.shannon <- ggplot(df.rptr, aes(rpt, variable)) +
  geom_point() +
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  xlab("Repeatability estimate") +
  ylab("") +
  scale_y_discrete(labels = c("Site", "Nest", "Bird ID", "Fixed")) +
  theme_classic()

rpt.confint.shannon
```

```{r}
paired.rpt.shannon <- cowplot::plot_grid(paired.boxplot.shannon, rpt.confint.shannon)

paired.rpt.shannon

#ggsave("Fig-_rpt.tiff", plot = paired.rpt, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 150)
```

# chao 
## Reaction norm plot

```{r}
ggplot(paired.df, aes(ageBinned, chao1, fill=ageBinned)) +
  geom_boxplot() +
  geom_line(aes(group=bird.ID), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  theme(legend.position = "none")
```

```{r}
paired.boxplot.chao <- ggplot(paired.df, aes(ageBinned, chao1, fill=ageBinned)) +
  geom_boxplot(aes(alpha = 0.1)) +
  geom_line(aes(group=bird.ID, colour = nest), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  xlab("Age") +
  ylab("Chao1 diversity") +
  theme_classic() +
  theme(legend.position = "none")

paired.boxplot.chao
```

## Plot w/ rptR estimate 

```{r}
df <- rpt.chao.birdID.nest.site.fixed

df.rptr <- cbind(t(df[["R"]]),df[["CI_emp"]]) %>% 
  tibble::rownames_to_column(var = "variable") %>%
  rename("rpt" = "t(df[[\"R\"]])", "lower" = "2.5%", "upper" = "97.5%") # including escapes for \"
  
df.rptr$variable <- factor(df.rptr$variable, ordered = T, levels = c("site", "nest", "bird.ID", "Fixed"))

df.rptr
```
I should add something here to indicate significance. Though is this small LRT values or CI's that dont overlap zero? Sometimes these disagree e.g. for site
```{r}
rpt.confint.chao <- ggplot(df.rptr, aes(rpt, variable)) +
  geom_point() +
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  xlab("Repeatability estimate") +
  ylab("") +
  scale_y_discrete(labels = c("Site", "Nest", "Bird ID", "Fixed")) +
  theme_classic()

rpt.confint.chao
```

```{r}
paired.rpt.chao <- cowplot::plot_grid(paired.boxplot.chao, rpt.confint.chao)

paired.rpt.chao

#ggsave("Fig-_rpt.tiff", plot = paired.rpt, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 150)
```


# Proteobacteria 
## Reaction norm plot


```{r}
ggplot(paired.df, aes(ageBinned, Proteobacteria.Abundance/NumberOfReads, fill=ageBinned)) +
  geom_boxplot() +
  geom_line(aes(group=bird.ID), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  theme(legend.position = "none")
```

```{r}
paired.boxplot <- ggplot(paired.df, aes(ageBinned, Proteobacteria.Abundance/NumberOfReads, fill=ageBinned)) +
  geom_boxplot(aes(alpha = 0.1)) +
  geom_line(aes(group=bird.ID, colour = nest), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  xlab("Age") +
  ylab("Proteobacteria Relative Abundance") +
  theme_classic() +
  theme(legend.position = "none")

paired.boxplot
```

## Plot w/ rptR estimate 

Note: there are two options for Proprtion data, 'R_org' or 'R_link'
* i will choose link here
* need drop= F to keep it in dataframe format for proper cbinding
```{r}
df <- rpt.pro.bird.nest.site.fixed

df.rptr <- cbind(t(df[["R"]])[,"R_link", drop = F], df[["CI_emp"]]$CI_link) %>% 
  tibble::rownames_to_column(var = "variable") %>%
  rename("rpt_link" = "R_link", "lower" = "2.5%", "upper" = "97.5%")
  
df.rptr$variable <- factor(df.rptr$variable, ordered = T, levels = c("site", "nest", "bird.ID", "Fixed"))

df.rptr
```

I should add something here to indicate significance. Though is this small LRT values or CI's that dont overlap zero? Sometimes these disagree e.g. for site
```{r}
rpt.confint.proteo <- ggplot(df.rptr, aes(rpt_link, variable)) +
  geom_point() +
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  xlab("Repeatability estimate") +
  ylab("") +
  scale_y_discrete(labels = c("Site", "Nest", "Bird ID", "Fixed")) +
  theme_classic()

rpt.confint.proteo
```

```{r}
paired.rpt.proteo <- cowplot::plot_grid(paired.boxplot, rpt.confint.proteo)

paired.rpt.proteo

#ggsave("Fig-_rpt.tiff", plot = paired.rpt, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 150)
```

# Firmicutes 
## Reaction norm plot


```{r}
ggplot(paired.df, aes(ageBinned, Firmicutes.Abundance/NumberOfReads, fill=ageBinned)) +
  geom_boxplot() +
  geom_line(aes(group=bird.ID), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  theme(legend.position = "none")
```

```{r}
paired.boxplot <- ggplot(paired.df, aes(ageBinned, Firmicutes.Abundance/NumberOfReads, fill=ageBinned)) +
  geom_boxplot(aes(alpha = 0.1)) +
  geom_line(aes(group=bird.ID, colour = nest), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  xlab("Age") +
  ylab("Firmicutes relative abundance") +
  theme_classic() +
  theme(legend.position = "none")

paired.boxplot
```

## Plot w/ rptR estimate 

```{r}
df <- rpt.fir.bird.nest.site.fixed

df.rptr <- cbind(t(df[["R"]])[,"R_link", drop = F], df[["CI_emp"]]$CI_link) %>% 
  tibble::rownames_to_column(var = "variable") %>%
  rename("rpt_link" = "R_link", "lower" = "2.5%", "upper" = "97.5%")
   
df.rptr$variable <- factor(df.rptr$variable, ordered = T, levels = c("site", "nest", "bird.ID", "Fixed"))

df.rptr
```
I should add something here to indicate significance. Though is this small LRT values or CI's that dont overlap zero? Sometimes these disagree e.g. for site
```{r}
rpt.confint.firmic <- ggplot(df.rptr, aes(rpt_link, variable)) +
  geom_point() +
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  xlab("Repeatability estimate") +
  ylab("") +
  scale_y_discrete(labels = c("Site", "Nest", "Bird ID", "Fixed")) +
  theme_classic()

rpt.confint.firmic
```

```{r}
paired.rpt.firmicutes <- cowplot::plot_grid(paired.boxplot, rpt.confint.firmic)

paired.rpt.firmicutes

#ggsave("Fig-_rpt.tiff", plot = paired.rpt, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 150)
```

# PC1 
## Reaction norm plot

```{r}
ggplot(paired.df, aes(ageBinned, PC1, fill=ageBinned)) +
  geom_boxplot() +
  geom_line(aes(group=bird.ID), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  theme(legend.position = "none")
```

```{r}
paired.boxplot <- ggplot(paired.df, aes(ageBinned, PC1, fill=ageBinned)) +
  geom_boxplot(aes(alpha = 0.1)) +
  geom_line(aes(group=bird.ID, colour = nest), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  xlab("Age") +
  ylab("PC1") +
  theme_classic() +
  theme(legend.position = "none")

paired.boxplot
```

## Plot w/ rptR estimate 

```{r}
df <- rpt.pc1.bird.nest.site.fixed

df.rptr <- cbind(t(df[["R"]]),df[["CI_emp"]]) %>% 
  tibble::rownames_to_column(var = "variable") %>%
  rename("rpt" = "t(df[[\"R\"]])", "lower" = "2.5%", "upper" = "97.5%") # including escapes for \"
  
df.rptr$variable <- factor(df.rptr$variable, ordered = T, levels = c("site", "nest", "bird.ID", "Fixed"))

df.rptr
```
I should add something here to indicate significance. Though is this small LRT values or CI's that dont overlap zero? Sometimes these disagree e.g. for site
```{r}
rpt.confint.pc1 <- ggplot(df.rptr, aes(rpt, variable)) +
  geom_point() +
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  xlab("Repeatability estimate") +
  ylab("") +
  scale_y_discrete(labels = c("Site", "Nest", "Bird ID", "Fixed")) +
  theme_classic()

rpt.confint.pc1
```

```{r}
paired.rpt.pc1 <- cowplot::plot_grid(paired.boxplot, rpt.confint.pc1)

paired.rpt.pc1

#ggsave("Fig-_rpt.tiff", plot = paired.rpt, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 150)
```

# PC2 
## Reaction norm plot


```{r}
ggplot(paired.df, aes(ageBinned, PC2, fill=ageBinned)) +
  geom_boxplot() +
  geom_line(aes(group=bird.ID), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  theme(legend.position = "none")
```

```{r}
paired.boxplot <- ggplot(paired.df, aes(ageBinned, PC2, fill=ageBinned)) +
  geom_boxplot(aes(alpha = 0.1)) +
  geom_line(aes(group=bird.ID, colour = nest), position = position_dodge(0.2)) +
  geom_point(aes(fill=ageBinned, group=bird.ID), position = position_dodge(0.2)) +
  xlab("Age") +
  ylab("PC2") +
  theme_classic() +
  theme(legend.position = "none")

paired.boxplot
```

## Plot w/ rptR estimate 

```{r}
df <- rpt.pc2.bird.nest.site.fixed

df.rptr <- cbind(t(df[["R"]]),df[["CI_emp"]]) %>% 
  tibble::rownames_to_column(var = "variable") %>%
  rename("rpt" = "t(df[[\"R\"]])", "lower" = "2.5%", "upper" = "97.5%") # including escapes for \"
  
df.rptr$variable <- factor(df.rptr$variable, ordered = T, levels = c("site", "nest", "bird.ID", "Fixed"))

df.rptr
```
I should add something here to indicate significance. Though is this small LRT values or CI's that dont overlap zero? Sometimes these disagree e.g. for site
```{r}
rpt.confint.pc2 <- ggplot(df.rptr, aes(rpt, variable)) +
  geom_point() +
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  xlab("Repeatability estimate") +
  ylab("") +
  scale_y_discrete(labels = c("Site", "Nest", "Bird ID", "Fixed")) +
  theme_classic()

rpt.confint.pc2
```

```{r}
paired.rpt.pc2 <- cowplot::plot_grid(paired.boxplot, rpt.confint.pc2)

paired.rpt.pc2

#ggsave("Fig-_rpt.tiff", plot = paired.rpt, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 150)
```

# Combine all plots

```{r}
paired.rpt.all <- cowplot::plot_grid(paired.rpt.shannon, paired.rpt.chao, paired.rpt.proteo, paired.rpt.firmicutes, paired.rpt.pc1, paired.rpt.pc2, ncol = 1)

paired.rpt.all
```

# repeatability plots only

```{r}
rpt.combine <- plot_grid(plot_grid(rpt.confint.shannon + ylab("Alpha \n diversity") + theme(axis.title.x=element_blank()),
                                  rpt.confint.chao+ theme(axis.title.x=element_blank()), 
                                                          nrow = 1, labels = c("A", "B")),
                         plot_grid(rpt.confint.proteo + ylab("Relative \n abundance")+ theme(axis.title.x=element_blank()),
                                   rpt.confint.firmic + theme(axis.title.x=element_blank()), 
                                   nrow = 1, labels = c("C", "D")),
                         plot_grid(rpt.confint.pc1 + ylab("Beta \n diversity"), rpt.confint.pc2, 
                                   nrow = 1, labels = c("E", "F")), nrow = 3)

rpt.combine
```

```{r}
ggsave("Fig6_rpt.tiff", plot = rpt.combine, device = "tiff", path = "../R_figures", width = 15, height = 10, units = "cm", dpi = 250)
```

