---
title: "R Notebook"
output: html_notebook
---

Understand results

# Standardize

What is happening with the habitat variable which following standardising is coded as (-0.793, 0.207) 

Run code in 03b_modelAveraging_AlphaDiv up to and including standardise chunk
* estimate for habitat is -0.9887422 in unstandardised model
* estimate is 0.04296 in standarised model
```{r}
summary(stdz.model)
```

```{r}
str(metadata)
```

```{r}
str(stdz.model@frame)
```

Does ageBinned keep its BDC contrasts? Yes, looks like it
```{r}
contrasts(stdz.model@frame$ageBinned)
```

```{r}
contrasts(metadata$habitat)
contrasts(stdz.model@frame$habitat)
```


###################################################################
Centring habitat
```{r}
metadata$habitat
habitat.binary <- ifelse(metadata$habitat=="deciduous",1,0 )

mean(habitat.binary) #!
```

###################################################################

```{r}
stdz.model <- standardize(all.sh.lmm, standardize.y = FALSE)#, binary.inputs = "0/1")#, unchanged = "habitat")#, binary.inputs = "0/1")
summary(stdz.model)

stdz.model@frame$c.habitat

table(stdz.model@frame$c.habitat)

```

```{r}
stdz.model <- standardize(all.sh.lmm, standardize.y = FALSE, unchanged = "habitat")#, binary.inputs = "0/1")#, unchanged = "habitat")#, binary.inputs = "0/1")
summary(stdz.model)

stdz.model@frame$habitat

#table(stdz.model@frame$c.habitat)

```

```{r}
stdz.model <- standardize(all.sh.lmm, standardize.y = FALSE, binary.inputs = "0/1")#, binary.inputs = "0/1")#, unchanged = "habitat")#, binary.inputs = "0/1")
summary(stdz.model)

stdz.model@frame$c.habitat

#table(stdz.model@frame$c.habitat)

```

##############################################
what does encoding habitat as factor do

```{r}
all.sh.lmm.factor <- all.sh.lmm
all.sh.lmm.factor@frame$habitat <- as.factor(all.sh.lmm.factor@frame$habitat)

str(all.sh.lmm.factor@frame$habitat)
```

```{r}
stdz.model <- standardize(all.sh.lmm.factor, standardize.y = FALSE)#, binary.inputs = "0/1")#, unchanged = "habitat")#, binary.inputs = "0/1")
summary(stdz.model)

stdz.model@frame$c.habitat

#table(stdz.model@frame$c.habitat)

str(stdz.model@frame)
```